---
title: "Project Viva: DNAm age performance and predictors"
author: 
  - name: "Anne Bozack"
output:
  html_document:
    toc: true
    toc_float: true
---


## Required Packages and functions
```{r,warning=FALSE,message=FALSE,eval=FALSE}
library(sas7bdat)
library(knitr)
library(dplyr)
library(kableExtra)
library(tidyverse)
# BiocManager::install("GAprediction")
library(psych)
library(Metrics)
library(GGally)
library(scales)
library(MASS)
library(lmtest)
library(sandwich)
```

## Load data
```{r,warning=FALSE,message=FALSE,eval=FALSE}
# Load data
# ageAll = read.csv('ageAll_DNAmAge_predictorsManuscript.csv')[,-1]

# Variables as factors
ageAll$smk_preg2 = factor(ageAll$smk_preg2)
ageAll$race_child2 = factor(ageAll$race_child2)
ageAll$female_d = as.factor(ageAll$female_d)
ageAll$coll_grad = as.factor(ageAll$coll_grad)
ageAll$gt70k = as.factor(ageAll$gt70k)
ageAll$preterm = factor(ageAll$preterm)

# Create dataset for each timepoint
datCord = ageAll[!is.na(ageAll$DNAmAge_cord),]
datAge3 = ageAll[!is.na(ageAll$DNAmAge_age3),]
datAge7 = ageAll[!is.na(ageAll$DNAmAge_age7),]
datAll = ageAll[!is.na(ageAll$DNAmAge_cord) & !is.na(ageAll$DNAmAge_age3) & !is.na(ageAll$DNAmAge_age7),]
```

## Descriptive statistics (Table 1 and Supplemental Table S1)
```{r,warning=FALSE,message=FALSE,eval=FALSE}
# function to calcuating descriptive stats
tableone = function(dat, contVars, factVars, title){
    rows = c(contVars, (unlist(sapply(1:length(factVars), function(i) c(factVars[i], levels(as.factor(dat[[factVars[i]]])))))))
    dfone = data.frame(matrix(ncol = 4, nrow = length(rows)))
    dfone[,1] = rows
    for (i in 1:length(contVars)){
        dfone[i,1] = paste0(dfone[i,1], ', mean, (SD)')
        dfone[i,2] = round(mean(dat[[contVars[i]]], na.rm = T), 2)
        dfone[i,3] = round(sd(dat[[contVars[i]]], na.rm = T), 2)
        dfone[i,4] = sum(is.na(dat[[contVars[i]]]))
    }
    rowN = length(contVars) + 1
    for (i in 1:length(factVars)){
        dfone[rowN,1] = paste0(dfone[rowN,1], ', n, (%)')
        rowN = rowN + 1
        for (j in 1:length(levels(as.factor(dat[[factVars[i]]])))){
            dfone[rowN,2] = table(as.factor(dat[[factVars[i]]]))[j]
            dfone[rowN,3] = round((table(as.factor(dat[[factVars[i]]]))/sum(!is.na(dat[[factVars[i]]])))[j]*100, 2)
            dfone[rowN,4] = sum(is.na(dat[[factVars[i]]]))
            rowN = rowN + 1
        }
    }
    dfone = rbind(c(paste0(title, ' N = ', nrow(dat)), NA, NA, NA), dfone)
    return(dfone)
}

contVars = c('age_mom_enroll_d', 'bmi_mom_prepreg_d', 'gestage_wks_deliv_d', 'zvalue_bwt_gage_sex_d')
factVars = c('coll_grad', 'smk_preg2', 'gt70k', 'female_d', 'race_child2', 'preterm')

tab1_cord = tableone(datCord, contVars, factVars, 'Cord blood')
tab1_age3 = tableone(datAge3, contVars, factVars, 'Age 3')
tab1_age7 = tableone(datAge7, contVars, factVars, 'Age 7')
tab1_all = tableone(datAll, contVars, factVars, 'All timepoints')

colnames(tab1_cord) = c('Var', '', '', 'Missing')
colnames(tab1_age3) = c('Var', '', '', 'Missing')
colnames(tab1_age7) = c('Var', '', '', 'Missing')
colnames(tab1_all) = c('Var', '', '', 'Missing')

tab1_cord[is.na(tab1_cord)] = ''
tab1_age3[is.na(tab1_age3)] = ''
tab1_age7[is.na(tab1_age7)] = ''
tab1_all[is.na(tab1_all)] = ''
```

### Cord blood
```{r, echo = F}
tab1_cord %>% kable() %>% kable_styling(font_size = 14) %>% scroll_box(width = "800px", height = "500px")
```

### Early childhood
```{r, echo = F}
tab1_age3 %>% kable() %>% kable_styling(font_size = 14) %>% scroll_box(width = "800px", height = "500px")
```

### Mid-childhood
```{r, echo = F}
tab1_age7 %>% kable() %>% kable_styling(font_size = 14) %>% scroll_box(width = "800px", height = "500px")
```

### Samples with data at all timepoints
```{r, echo = F}
tab1_all %>% kable() %>% kable_styling(font_size = 14) %>% scroll_box(width = "800px", height = "500px")
```

## Correlations with chronological age and among clocks (Figures 2 and 3)
```{r,warning=FALSE,message=FALSE}
# Cord blood GA and EGA
cordEGA_cor = corr.test(datCord[,c('gestage_wks_deliv_d', 'GA_BohlinPred_wks', 'GA_KnightPred_wks')])

# Cord blood GA and EA
cordEA_cor = corr.test(datCord[,c('GA_wks_adj39', 'DNAmAge_cord_wks', 'DNAmAgeSkinBloodClock_cord_wks')])

# Early childhood age and EA
age3_cor = corr.test(datAge3[,c('Age_age3', 'DNAmAge_age3', 'DNAmAgeSkinBloodClock_age3')])

# Mid-childhood age and EA
age7_cor = corr.test(datAge7[,c('Age_age7', 'DNAmAge_age7', 'DNAmAgeSkinBloodClock_age7')])
```

### Cord blood and EGA
#### Correlation coefficients
```{r, echo = F}
cordEGA_cor$r %>% kable() %>% kable_styling(font_size = 14) 
```

#### p-values
```{r, echo = F}
cordEGA_cor$p %>% kable() %>% kable_styling(font_size = 14) 
```

### Cord blood and EA
#### Correlation coefficients
```{r, echo = F}
cordEA_cor$r %>% kable() %>% kable_styling(font_size = 14) 
```

#### p-values
```{r, echo = F}
cordEA_cor$p %>% kable() %>% kable_styling(font_size = 14) 
```

### Early childhood
#### Correlation coefficients
```{r, echo = F}
age3_cor$r %>% kable() %>% kable_styling(font_size = 14) 
```

#### p-values
```{r, echo = F}
age3_cor$p %>% kable() %>% kable_styling(font_size = 14) 
```

### Mid-childhood
#### Correlation coefficients
```{r, echo = F}
age7_cor$r %>% kable() %>% kable_styling(font_size = 14) 
```

#### p-values
```{r, echo = F}
age7_cor$p %>% kable() %>% kable_styling(font_size = 14) 
```


## MAE with chronological age (shown in Figures 2 and 3)
```{r,warning=FALSE,message=FALSE}
# Cord blood and EGA
cordMAE_EGA = data.frame(matrix(nrow = 1, ncol = 2))
colnames(cordMAE_EGA) = c('Bohlin', 'Knight')
rownames(cordMAE_EGA) = 'Chronological GA'
cordMAE_EGA[1,1] = mdae(datCord$gestage_wks_deliv_d, datCord$GA_BohlinPred_wks)
cordMAE_EGA[1,2] = mdae(datCord$gestage_wks_deliv_d, datCord$GA_KnightPred_wks)

# Cord blood and EA (using GA converted to years [(gestational week – 39)/52])
cordMAE_EA = data.frame(matrix(nrow = 1, ncol = 2))
colnames(cordMAE_EA) = c('Horvath', 'Skinblood')
rownames(cordMAE_EA) = 'Chronological GA'
cordMAE_EA[1,1] = mdae(datCord$GA_wks_adj39, datCord$DNAmAge_cord_wks)
cordMAE_EA[1,2] = mdae(datCord$GA_wks_adj39, datCord$DNAmAgeSkinBloodClock_cord_wks)

# Early childhood
age3MAE = data.frame(matrix(nrow = 1, ncol = 2))
colnames(age3MAE) = c('Horvath', 'Skinblood')
rownames(age3MAE) = 'Chronological age'
age3MAE[1,1] = mdae(datAge3$Age_age3, datAge3$DNAmAge_age3)
age3MAE[1,2] = mdae(datAge3$Age_age3, datAge3$DNAmAgeSkinBloodClock_age3)

# Early childhood
age7MAE = data.frame(matrix(nrow = 1, ncol = 2))
colnames(age7MAE) = c('Horvath', 'Skinblood')
rownames(age7MAE) = 'Chronological age'
age7MAE[1,1] = mdae(datAge7$Age_age7, datAge7$DNAmAge_age7)
age7MAE[1,2] = mdae(datAge7$Age_age7, datAge7$DNAmAgeSkinBloodClock_age7)
```

### Cord blood and EGA
```{r, echo = F}
cordMAE_EGA %>% kable() %>% kable_styling(font_size = 14) 
```

### Cord blood and EA
```{r, echo = F}
cordMAE_EA %>% kable() %>% kable_styling(font_size = 14) 
```

### Early childhood
```{r, echo = F}
age3MAE %>% kable() %>% kable_styling(font_size = 14) 
```

### Mid-childhood
```{r, echo = F}
age7MAE %>% kable() %>% kable_styling(font_size = 14) 
```

## Adjusting for GA in early and mid-childhood estimates (Supplemental Figure 1)
```{r,warning=FALSE,message=FALSE}
datAge3$Age_age3_GAadj = datAge3$Age_age3 + datAge3$gestage_wks_deliv_d/52
datAge7$Age_age7_GAadj = datAge7$Age_age7 + datAge7$gestage_wks_deliv_d/52

datAge3$DNAm_absDiff = abs(datAge3$DNAmAge_age3 - datAge3$Age_age3)
datAge7$DNAm_absDiff = abs(datAge7$DNAmAge_age7 - datAge7$Age_age7)
datAge3$DNAm_absDiff_GAadj = abs(datAge3$DNAmAge_age3 - datAge3$Age_age3_GAadj)
datAge7$DNAm_absDiff_GAadj = abs(datAge7$DNAmAge_age7 - datAge7$Age_age7_GAadj)

# MAE after adjusting for GA
mdae(datAge3$Age_age3_GAadj, datAge3$DNAmAge_age3)
mdae(datAge7$Age_age7_GAadj, datAge7$DNAmAge_age7)

# test for differences with and without GA adjustment
wilcox.test(datAge3$DNAm_absDiff, datAge3$DNAm_absDiff_GAadj, paired = T)

wilcox.test(datAge7$DNAm_absDiff, datAge7$DNAm_absDiff_GAadj, paired = T)
```

## Spearman correlations with cell type (Figure 4)
```{r,warning=FALSE,message=FALSE,echo= FALSE}
# Cord blood
cellCord = c("Bcell_perc_cord", "CD4T_perc_cord", "CD8T_perc_cord", "Gran_perc_cord", "Mono_perc_cord", "NK_perc_cord", "nRBC_perc_cord")

cell_cord_df = data.frame(y = rep(c('% B cells', '% CD4+ T cells', '% CD8+ T cells', '% granulocytes', '% monocytes', '% NK cells', '% nucleated RBCs'), times = 5),
    x = rep(c('Chronological GA  ', 'Bohlin  EGAA  ', 'Knight  EGAA  ', 'Horvath EAA ', 'Skin & blood EAA   '), each = 7), r = NA, p = NA)

agevars = c('gestage_wks_deliv_d', 'GA_BohlinPred_EAA_wks', 'GA_KnightPred_EAA_wks', 'AgeAccelerationResidual_cord_wks', 'DNAmAgeSkinBloodClockAdjAge_cord_wks')
row = 1
for (i in 1:length(agevars)){
    for (j in 1:length(cellCord)){
        cell_cord_df[row,3] = cor.test(datCord[[cellCord[j]]], datCord[[agevars[[i]]]], method = 'spearman', exact = F)[[4]]
        cell_cord_df[row,4] = cor.test(datCord[[cellCord[j]]], datCord[[agevars[[i]]]], method = 'spearman', exact = F)[[3]]
        row = row + 1
    }
}

# Early childhood
cellAge3 = c("Bcell_perc_age3", "CD4T_perc_age3", "CD8T_perc_age3", "Neu_perc_age3", "Mono_perc_age3", "NK_perc_age3")

cell_age3_df = data.frame(y = rep(c('% B cells', '% CD4+ T cells', '% CD8+ T cells', '% neutrophils', '% monocytes', '% NK cells'), times = 3),
    x = rep(c('Chronological\n age ', 'Horvath\n EAA ', 'Skin & blood\n EAA   '), each = 6), r = NA, p = NA)

agevars = c('Age_age3', 'AgeAccelerationResidual_age3', 'DNAmAgeSkinBloodClockAdjAge_age3')
row = 1
for (i in 1:length(agevars)){
    for (j in 1:length(cellAge3)){
        cell_age3_df[row,3] = cor.test(datAge3[[cellAge3[j]]], datAge3[[agevars[[i]]]], method = 'spearman', exact = F)[[4]]
        cell_age3_df[row,4] = cor.test(datAge3[[cellAge3[j]]], datAge3[[agevars[[i]]]], method = 'spearman', exact = F)[[3]]
        row = row + 1
    }
}

# Mid-childhood
cellAge7 = c("Bcell_perc_age7", "CD4T_perc_age7", "CD8T_perc_age7", "Neu_perc_age7", "Mono_perc_age7", "NK_perc_age7")

cell_age7_df = data.frame(y = rep(c('% B cells', '% CD4+ T cells', '% CD8+ T cells', '% neutrophils', '% monocytes', '% NK cells'), times = 3),
    x = rep(c('Chronological\n age ', 'Horvath\n EAA ', 'Skin & blood\n EAA   '), each = 6), r = NA, p = NA)

agevars = c('Age_age7', 'AgeAccelerationResidual_age7', 'DNAmAgeSkinBloodClockAdjAge_age7')
row = 1
for (i in 1:length(agevars)){
    for (j in 1:length(cellAge7)){
        cell_age7_df[row,3] = cor.test(datAge7[[cellAge7[j]]], datAge7[[agevars[[i]]]], method = 'spearman', exact = F)[[4]]
        cell_age7_df[row,4] = cor.test(datAge7[[cellAge7[j]]], datAge7[[agevars[[i]]]], method = 'spearman', exact = F)[[3]]
        row = row + 1
    }
}
```

### Cord blood
```{r, echo = F}
cell_cord_df %>% kable() %>% kable_styling(font_size = 14) 
```

### Early childhood
```{r, echo = F}
cell_age3_df %>% kable() %>% kable_styling(font_size = 14) 
```

### Mid-childhood
```{r, echo = F}
cell_age7_df %>% kable() %>% kable_styling(font_size = 14) 
```

## Associations with maternal-infant characteristics
### Cord blood (Figure 5, Supplemental Table 2)
```{r,warning=FALSE,message=FALSE,eval= FALSE}
# Bohlin
mod_multivar_cord_Bohlin = rlm(GA_BohlinPred_EAA_wks ~ female_d + preterm + zvalue_bwt_gage_sex_d + race_child2 + agemom_scale + bmi_scale + coll_grad + smk_preg2, data = datCord)
coef = coeftest(mod_multivar_cord_Bohlin, vcovHC(mod_multivar_cord_Bohlin, type="HC0"))
ci = coefci(x = mod_multivar_cord_Bohlin, level = 0.95, vcov. = vcovHC(mod_multivar_cord_Bohlin, type="HC0"))
multivar_cord_Bohlin = data.frame(B_boh = coef[,1], CI_boh_low = ci[,1], CI_boh_high = ci[,2], summary_boh = paste0(round(coef[,1], 2), ' (', round(ci[,1], 2), ", ", round(ci[,2], 2), ')'), p_boh = round(coef[,4],3))

mod_multivar_cell_cord_Bohlin = rlm(GA_BohlinPred_EAA_wks ~ female_d + preterm + zvalue_bwt_gage_sex_d + race_child2 + agemom_scale + bmi_scale + coll_grad + smk_preg2 + Bcell_perc_cord + CD4T_perc_cord + CD8T_perc_cord + Gran_perc_cord + Mono_perc_cord + NK_perc_cord + nRBC_perc_cord, data = datCord)
coef = coeftest(mod_multivar_cell_cord_Bohlin, vcovHC(mod_multivar_cell_cord_Bohlin, type="HC0"))
ci = coefci(x = mod_multivar_cell_cord_Bohlin, level = 0.95, vcov. = vcovHC(mod_multivar_cell_cord_Bohlin, type="HC0"))
multivar_cell_cord_Bohlin = data.frame(B_boh_cellAdj = coef[,1], CI_boh_low_cellAdj = ci[,1], CI_boh_high_cellAdj = ci[,2], summary_boh_cellAdj = paste0(round(coef[,1], 2), ' (', round(ci[,1], 2), ", ", round(ci[,2], 2), ')'), p_boh_cellAdj = round(coef[,4], 3))

# Knight
mod_multivar_cord_Knight = rlm(GA_KnightPred_EAA_wks ~ female_d + preterm + zvalue_bwt_gage_sex_d + race_child2 + agemom_scale + bmi_scale + coll_grad + smk_preg2, data = datCord)
coef = coeftest(mod_multivar_cord_Knight, vcovHC(mod_multivar_cord_Knight, type="HC0"))
ci = coefci(x = mod_multivar_cord_Knight, level = 0.95, vcov. = vcovHC(mod_multivar_cord_Knight, type="HC0"))
multivar_cord_Knight = data.frame(B_kn = coef[,1], CI_kn_low = ci[,1], CI_kn_high = ci[,2], summary_kn = paste0(round(coef[,1], 2), ' (', round(ci[,1], 2), ", ", round(ci[,2], 2), ')'), p_kn = round(coef[,4], 3))

mod_multivar_cell_cord_Knight = rlm(GA_KnightPred_EAA_wks ~ female_d + preterm + zvalue_bwt_gage_sex_d + race_child2 + agemom_scale + bmi_scale + coll_grad + smk_preg2 + Bcell_perc_cord + CD4T_perc_cord + CD8T_perc_cord + Gran_perc_cord + Mono_perc_cord + NK_perc_cord + nRBC_perc_cord, data = datCord)
coef = coeftest(mod_multivar_cell_cord_Knight, vcovHC(mod_multivar_cell_cord_Knight, type="HC0"))
ci = coefci(x = mod_multivar_cell_cord_Knight, level = 0.95, vcov. = vcovHC(mod_multivar_cell_cord_Knight, type="HC0"))
multivar_cell_cord_Knight = data.frame(B_kn_cellAdj = coef[,1], CI_kn_low_cellAdj = ci[,1], CI_kn_high_cellAdj = ci[,2], summary_kn_cellAdj = paste0(round(coef[,1], 2), ' (', round(ci[,1], 2), ", ", round(ci[,2], 2), ')'), p_kn_cellAdj = round(coef[,4], 3))

# Horvath
mod_multivar_cord_Horvath = rlm(AgeAccelerationResidual_cord_wks ~ female_d + preterm + zvalue_bwt_gage_sex_d + race_child2 + agemom_scale + bmi_scale + coll_grad + smk_preg2, data = datCord)
coef = coeftest(mod_multivar_cord_Horvath, vcovHC(mod_multivar_cord_Horvath, type="HC0"))
ci = coefci(x = mod_multivar_cord_Horvath, level = 0.95, vcov. = vcovHC(mod_multivar_cord_Horvath, type="HC0"))
multivar_cord_Horvath = data.frame(B_hor = coef[,1], CI_hor_low = ci[,1], CI_hor_high = ci[,2], summary_hor = paste0(round(coef[,1], 2), ' (', round(ci[,1], 2), ", ", round(ci[,2], 2), ')'), p_hor = round(coef[,4], 3))

mod_multivar_cell_cord_Horvath = rlm(AgeAccelerationResidual_cord_wks ~ female_d + preterm + zvalue_bwt_gage_sex_d + race_child2 + agemom_scale + bmi_scale + coll_grad + smk_preg2 + Bcell_perc_cord + CD4T_perc_cord + CD8T_perc_cord + Gran_perc_cord + Mono_perc_cord + NK_perc_cord + nRBC_perc_cord, data = datCord)
coef = coeftest(mod_multivar_cell_cord_Horvath, vcovHC(mod_multivar_cell_cord_Horvath, type="HC0"))
ci = coefci(x = mod_multivar_cell_cord_Horvath, level = 0.95, vcov. = vcovHC(mod_multivar_cell_cord_Horvath, type="HC0"))
multivar_cord_Horvath = data.frame(B_hor_cellAdj = coef[,1], CI_hor_low_cellAdj = ci[,1], CI_hor_high_cellAdj = ci[,2], summary_hor_cellAdj = paste0(round(coef[,1], 2), ' (', round(ci[,1], 2), ", ", round(ci[,2], 2), ')'), p_hor_cellAdj = round(coef[,4], 3))

# skin & blood
mod_multivar_cord_skinblood = rlm(DNAmAgeSkinBloodClockAdjAge_cord_wks ~ female_d + preterm + zvalue_bwt_gage_sex_d + race_child2 + agemom_scale + bmi_scale + coll_grad + smk_preg2, data = datCord)
coef = coeftest(mod_multivar_cord_skinblood, vcovHC(mod_multivar_cord_skinblood, type="HC0"))
ci = coefci(x = mod_multivar_cord_skinblood, level = 0.95, vcov. = vcovHC(mod_multivar_cord_skinblood, type="HC0"))
multivar_cord_skinblood = data.frame(B_sb = coef[,1], CI_sb_low = ci[,1], CI_sb_high = ci[,2], summary_sb = paste0(round(coef[,1], 2), ' (', round(ci[,1], 2), ", ", round(ci[,2], 2), ')'), p_sb = round(coef[,4], 3))

mod_multivar_cell_cord_skinblood = rlm(DNAmAgeSkinBloodClockAdjAge_cord_wks ~ female_d + preterm + zvalue_bwt_gage_sex_d + race_child2 + agemom_scale + bmi_scale + coll_grad + smk_preg2 + Bcell_perc_cord + CD4T_perc_cord + CD8T_perc_cord + Gran_perc_cord + Mono_perc_cord + NK_perc_cord + nRBC_perc_cord, data = datCord)
coef = coeftest(mod_multivar_cell_cord_skinblood, vcovHC(mod_multivar_cell_cord_skinblood, type="HC0"))
ci = coefci(x = mod_multivar_cell_cord_skinblood, level = 0.95, vcov. = vcovHC(mod_multivar_cell_cord_skinblood, type="HC0"))
multivar_cell_cord_skinblood = data.frame(B_sb_cellAdj = coef[,1], CI_sb_low_cellAdj = ci[,1], CI_sb_high_cellAdj = ci[,2], summary_sb_cellAdj = paste0(round(coef[,1], 2), ' (', round(ci[,1], 2), ", ", round(ci[,2], 2), ')'), p_sb_cellAdj = round(coef[,4], 3))
```

#### Bohlin EAA
```{r, echo = F}
cbind(multivar_cord_Bohlin[-1,c(4,5)], multivar_cell_cord_Bohlin[c(2:13),c(4,5)]) %>% kable() %>% kable_styling(font_size = 14) %>% add_header_above(c(" " = 1, "Not adjusted for cell type" = 2, "Adjusted for cellt type" = 2))
```

#### Knight EAA
```{r, echo = F}
cbind(multivar_cord_Knight[-1,c(4,5)], multivar_cell_cord_Knight[c(2:13),c(4,5)]) %>% kable() %>% kable_styling(font_size = 14) %>% add_header_above(c(" " = 1, "Not adjusted for cell type" = 2, "Adjusted for cellt type" = 2))
```

#### Horvath EA
```{r, echo = F}
cbind(multivar_cord_Horvath[-1,c(4,5)], multivar_cell_cord_Horvath[c(2:13),c(4,5)]) %>% kable() %>% kable_styling(font_size = 14) %>% add_header_above(c(" " = 1, "Not adjusted for cell type" = 2, "Adjusted for cellt type" = 2))
```

#### Skin & blood EA
```{r, echo = F}
cbind(multivar_cord_skinblood[-1,c(4,5)], multivar_cell_cord_skinblood[c(2:13),c(4,5)]) %>% kable() %>% kable_styling(font_size = 14) %>% add_header_above(c(" " = 1, "Not adjusted for cell type" = 2, "Adjusted for cellt type" = 2))
```

### Early and mid-childhood (Figure 6, Supplemental Table 3)
```{r,warning=FALSE,message=FALSE,eval= FALSE}
# Early childhood Horvath
mod_multivar_age3_Horvath = rlm(AgeAccelerationResidual_age3 ~ female_d + preterm + zvalue_bwt_gage_sex_d + race_child2 + agemom_scale + bmi_scale + coll_grad + smk_preg2, data = datAge3)
coef = coeftest(mod_multivar_age3_Horvath, vcovHC(mod_multivar_age3_Horvath, type="HC0"))
ci = coefci(x = mod_multivar_age3_Horvath, level = 0.95, vcov. = vcovHC(mod_multivar_age3_Horvath, type="HC0"))
multivar_age3_Horvath = data.frame(B_age3Hor = coef[,1], CI_age3Hor_low = ci[,1], CI_age3Hor_high = ci[,2], summary_age3Hor = paste0(round(coef[,1], 2), ' (', round(ci[,1], 2), ", ", round(ci[,2], 2), ')'), p_boh = round(coef[,4],3))

mod_multivar_cell_age3_Horvath = rlm(AgeAccelerationResidual_age3 ~ female_d + preterm + zvalue_bwt_gage_sex_d + race_child2 + agemom_scale + bmi_scale + coll_grad + smk_preg2 + Bcell_perc_age3 + CD4T_perc_age3 + CD8T_perc_age3 + Neu_perc_age3 + Mono_perc_age3 + NK_perc_age3, data = datAge3)
coef = coeftest(mod_multivar_cell_age3_Horvath, vcovHC(mod_multivar_cell_age3_Horvath, type="HC0"))
ci = coefci(x = mod_multivar_cell_age3_Horvath, level = 0.95, vcov. = vcovHC(mod_multivar_cell_age3_Horvath, type="HC0"))
multivar_cell_age3_Horvath = data.frame(B_age3Hor_cellAdj = coef[,1], CI_age3Hor_low_cellAdj = ci[,1], CI_age3Hor_high_cellAdj = ci[,2], summary_age3Hor_cellAdj = paste0(round(coef[,1], 2), ' (', round(ci[,1], 2), ", ", round(ci[,2], 2), ')'), p_boh_cellAdj = round(coef[,4], 3))

# Early childhood skin & blood
mod_multivar_age3_skinblood = rlm(DNAmAgeSkinBloodClockAdjAge_age3 ~ female_d + preterm + zvalue_bwt_gage_sex_d + race_child2 + agemom_scale + bmi_scale + coll_grad + smk_preg2, data = datAge3)
coef = coeftest(mod_multivar_age3_skinblood, vcovHC(mod_multivar_age3_skinblood, type="HC0"))
ci = coefci(x = mod_multivar_age3_skinblood, level = 0.95, vcov. = vcovHC(mod_multivar_age3_skinblood, type="HC0"))
multivar_age3_skinblood = data.frame(B_age3skin = coef[,1], CI_age3skin_low = ci[,1], CI_age3skin_high = ci[,2], summary_age3skin = paste0(round(coef[,1], 2), ' (', round(ci[,1], 2), ", ", round(ci[,2], 2), ')'), p_kn = round(coef[,4], 3))

mod_multivar_cell_age3_skinblood = rlm(DNAmAgeSkinBloodClockAdjAge_age3 ~ female_d + preterm + zvalue_bwt_gage_sex_d + race_child2 + agemom_scale + bmi_scale + coll_grad + smk_preg2 + Bcell_perc_age3 + CD4T_perc_age3 + CD8T_perc_age3 + Neu_perc_age3 + Mono_perc_age3 + NK_perc_age3, data = datAge3)
coef = coeftest(mod_multivar_cell_age3_skinblood, vcovHC(mod_multivar_cell_age3_skinblood, type="HC0"))
ci = coefci(x = mod_multivar_cell_age3_skinblood, level = 0.95, vcov. = vcovHC(mod_multivar_cell_age3_skinblood, type="HC0"))
multivar_cell_age3_skinblood = data.frame(B_age3skin_cellAdj = coef[,1], CI_age3skin_low_cellAdj = ci[,1], CI_age3skin_high_cellAdj = ci[,2], summary_age3skin_cellAdj = paste0(round(coef[,1], 2), ' (', round(ci[,1], 2), ", ", round(ci[,2], 2), ')'), p_kn_cellAdj = round(coef[,4], 3))

# Mid-chiildhood Horvath
mod_multivar_age7_Horvath = rlm(AgeAccelerationResidual_age7 ~ female_d + preterm + zvalue_bwt_gage_sex_d + race_child2 + agemom_scale + bmi_scale + coll_grad + smk_preg2, data = datAge7)
coef = coeftest(mod_multivar_age7_Horvath, vcovHC(mod_multivar_age7_Horvath, type="HC0"))
ci = coefci(x = mod_multivar_age7_Horvath, level = 0.95, vcov. = vcovHC(mod_multivar_age7_Horvath, type="HC0"))
multivar_age7_Horvath = data.frame(B_age7Hor = coef[,1], CI_age7Hor_low = ci[,1], CI_age7Hor_high = ci[,2], summary_age7Hor = paste0(round(coef[,1], 2), ' (', round(ci[,1], 2), ", ", round(ci[,2], 2), ')'), p_boh = round(coef[,4],3))

mod_multivar_cell_age7_Horvath = rlm(AgeAccelerationResidual_age7 ~ female_d + preterm + zvalue_bwt_gage_sex_d + race_child2 + agemom_scale + bmi_scale + coll_grad + smk_preg2 + Bcell_perc_age7 + CD4T_perc_age7 + CD8T_perc_age7 + Neu_perc_age7 + Mono_perc_age7 + NK_perc_age7, data = datAge7)
coef = coeftest(mod_multivar_cell_age7_Horvath, vcovHC(mod_multivar_cell_age7_Horvath, type="HC0"))
ci = coefci(x = mod_multivar_cell_age7_Horvath, level = 0.95, vcov. = vcovHC(mod_multivar_cell_age7_Horvath, type="HC0"))
multivar_cell_age7_Horvath = data.frame(B_age7Hor_cellAdj = coef[,1], CI_age7Hor_low_cellAdj = ci[,1], CI_age7Hor_high_cellAdj = ci[,2], summary_age7Hor_cellAdj = paste0(round(coef[,1], 2), ' (', round(ci[,1], 2), ", ", round(ci[,2], 2), ')'), p_boh_cellAdj = round(coef[,4], 3))

# Mid-childhood skin & blood
mod_multivar_age7_skinblood = rlm(DNAmAgeSkinBloodClockAdjAge_age7 ~ female_d + preterm + zvalue_bwt_gage_sex_d + race_child2 + agemom_scale + bmi_scale + coll_grad + smk_preg2, data = datAge7)
coef = coeftest(mod_multivar_age7_skinblood, vcovHC(mod_multivar_age7_skinblood, type="HC0"))
ci = coefci(x = mod_multivar_age7_skinblood, level = 0.95, vcov. = vcovHC(mod_multivar_age7_skinblood, type="HC0"))
multivar_age7_skinblood = data.frame(B_age7skin = coef[,1], CI_age7skin_low = ci[,1], CI_age7skin_high = ci[,2], summary_age7skin = paste0(round(coef[,1], 2), ' (', round(ci[,1], 2), ", ", round(ci[,2], 2), ')'), p_kn = round(coef[,4], 3))

mod_multivar_cell_age7_skinblood = rlm(DNAmAgeSkinBloodClockAdjAge_age7 ~ female_d + preterm + zvalue_bwt_gage_sex_d + race_child2 + agemom_scale + bmi_scale + coll_grad + smk_preg2 + Bcell_perc_age7 + CD4T_perc_age7 + CD8T_perc_age7 + Neu_perc_age7 + Mono_perc_age7 + NK_perc_age7, data = datAge7)
coef = coeftest(mod_multivar_cell_age7_skinblood, vcovHC(mod_multivar_cell_age7_skinblood, type="HC0"))
ci = coefci(x = mod_multivar_cell_age7_skinblood, level = 0.95, vcov. = vcovHC(mod_multivar_cell_age7_skinblood, type="HC0"))
multivar_cell_age7_skinblood = data.frame(B_age7skin_cellAdj = coef[,1], CI_age7skin_low_cellAdj = ci[,1], CI_age7skin_high_cellAdj = ci[,2], summary_age7skin_cellAdj = paste0(round(coef[,1], 2), ' (', round(ci[,1], 2), ", ", round(ci[,2], 2), ')'), p_kn_cellAdj = round(coef[,4], 3))
```

#### Early childhood Horvath EA
```{r, echo = F}
cbind(multivar_age3_Horvath[-1,c(4,5)], multivar_cell_age3_Horvath[c(2:13),c(4,5)]) %>% kable() %>% kable_styling(font_size = 14) %>% add_header_above(c(" " = 1, "Not adjusted for cell type" = 2, "Adjusted for cellt type" = 2))
```

#### Early childhood skin & blood EA
```{r, echo = F}
cbind(multivar_age3_skinblood[-1,c(4,5)], multivar_cell_age3_skinblood[c(2:13),c(4,5)]) %>% kable() %>% kable_styling(font_size = 14) %>% add_header_above(c(" " = 1, "Not adjusted for cell type" = 2, "Adjusted for cellt type" = 2))
```

#### Mid-childhood Horvath EA
```{r, echo = F}
cbind(multivar_age7_Horvath[-1,c(4,5)], multivar_cell_age7_Horvath[c(2:13),c(4,5)]) %>% kable() %>% kable_styling(font_size = 14) %>% add_header_above(c(" " = 1, "Not adjusted for cell type" = 2, "Adjusted for cellt type" = 2))
```

#### Mid-childhood skin & blood EA
```{r, echo = F}
cbind(multivar_age7_skinblood[-1,c(4,5)], multivar_cell_age7_skinblood[c(2:13),c(4,5)]) %>% kable() %>% kable_styling(font_size = 14) %>% add_header_above(c(" " = 1, "Not adjusted for cell type" = 2, "Adjusted for cellt type" = 2))
```

## Associations between cord blood Horvath EA and childhoood EAA (Figure 7, Supplemental Table 4)
```{r,warning=FALSE,message=FALSE,echo= FALSE, eval = F}
# Early childhood
# dataframe to store results
EAassoc_age3 = data.frame(matrix(nrow = 2, ncol = 6))
colnames(EAassoc_age3) = c('Horvath_B', 'Horvath_CI', 'Horvath_p', 'age3_skinblood_B', 'age3_skinblood_CI', 'age3_skinblood_p')
rownames(EAassoc_age3) = c('Unadjusted', 'Adjusted')

# Horvath EAA unadjusted
horvathAge3Fit = rlm(AgeAccelerationResidual_age3 ~ DNAmAge_cord_wks, data = datAge3)
EAassoc_age3[1,1] = coeftest(horvathAge3Fit, vcovHC(horvathAge3Fit, type="HC0"))[2,1]
EAassoc_age3[1,2] = paste0('(', round(coefci(x = horvathAge3Fit, level = 0.95, vcov. = vcovHC(horvathAge3Fit, type="HC0"))[2,1], 3), ', ', round(coefci(x = horvathAge3Fit, level = 0.95, vcov. = vcovHC(horvathAge3Fit, type="HC0"))[2,2], 3))
EAassoc_age3[1,3] = coeftest(horvathAge3Fit, vcovHC(horvathAge3Fit, type="HC0"))[2,4]

# Skin & bood EAA unadjusted
skinbloodAge3Fit = rlm(DNAmAgeSkinBloodClockAdjAge_age3 ~ DNAmAge_cord_wks, data = datAge3)
EAassoc_age3[1,4] = coeftest(skinbloodAge3Fit, vcovHC(skinbloodAge3Fit, type="HC0"))[2,1]
EAassoc_age3[1,5] = paste0('(', round(coefci(x = skinbloodAge3Fit, level = 0.95, vcov. = vcovHC(skinbloodAge3Fit, type="HC0"))[2,1], 3), ', ', round(coefci(x = skinbloodAge3Fit, level = 0.95, vcov. = vcovHC(skinbloodAge3Fit, type="HC0"))[2,2], 3))
EAassoc_age3[1,6] = coeftest(skinbloodAge3Fit, vcovHC(skinbloodAge3Fit, type="HC0"))[2,4]

# Horvath EAA adjusted
horvathAge3FitAdj = rlm(AgeAccelerationResidual_age3 ~ DNAmAge_cord_wks + female_d + preterm + zvalue_bwt_gage_sex_d + race_child2 + agemom_scale + bmi_scale + coll_grad + smk_preg2 + Bcell_perc_age3 + CD4T_perc_age3 + CD8T_perc_age3 + Neu_perc_age3 + Mono_perc_age3 + NK_perc_age3, data = datAge3)
EAassoc_age3[2,1] = coeftest(horvathAge3FitAdj, vcovHC(horvathAge3FitAdj, type="HC0"))[2,1]
EAassoc_age3[2,2] = paste0('(', round(coefci(x = horvathAge3FitAdj, level = 0.95, vcov. = vcovHC(horvathAge3FitAdj, type="HC0"))[2,1], 3), ', ', round(coefci(x = horvathAge3FitAdj, level = 0.95, vcov. = vcovHC(horvathAge3FitAdj, type="HC0"))[2,2], 3))
EAassoc_age3[2,3] = coeftest(horvathAge3FitAdj, vcovHC(horvathAge3FitAdj, type="HC0"))[2,4]

# Skin & blood EAA adjusted
skinbloodAge3FitAdj = rlm(DNAmAgeSkinBloodClockAdjAge_age3 ~ DNAmAge_cord_wks + female_d + preterm + zvalue_bwt_gage_sex_d + race_child2 + agemom_scale + bmi_scale + coll_grad + smk_preg2 + Bcell_perc_age3 + CD4T_perc_age3 + CD8T_perc_age3 + Neu_perc_age3 + Mono_perc_age3 + NK_perc_age3, data = datAge3)
EAassoc_age3[2,4] = coeftest(skinbloodAge3FitAdj, vcovHC(skinbloodAge3FitAdj, type="HC0"))[2,1]
EAassoc_age3[2,5] = paste0('(', round(coefci(x = skinbloodAge3FitAdj, level = 0.95, vcov. = vcovHC(skinbloodAge3FitAdj, type="HC0"))[2,1], 3), ', ', round(coefci(x = skinbloodAge3FitAdj, level = 0.95, vcov. = vcovHC(skinbloodAge3FitAdj, type="HC0"))[2,2], 3))
EAassoc_age3[2,6] = coeftest(skinbloodAge3FitAdj, vcovHC(skinbloodAge3FitAdj, type="HC0"))[2,4]


# Mid-childhood
# dataframe to store results
EAassoc_age7 = data.frame(matrix(nrow = 2, ncol = 6))
colnames(EAassoc_age7) = c('Horvath_B', 'Horvath_CI', 'Horvath_p', 'age3_skinblood_B', 'age3_skinblood_CI', 'age3_skinblood_p')
rownames(EAassoc_age7) = c('Unadjusted', 'Adjusted')

# Horvath EAA =unadjusted
horvathAge7Fit = rlm(AgeAccelerationResidual_age7 ~ DNAmAge_cord_wks, data = datAge7)
EAassoc_age7[1,1] = coeftest(horvathAge7Fit, vcovHC(horvathAge7Fit, type="HC0"))[2,1]
EAassoc_age7[1,2] = paste0('(', round(coefci(x = horvathAge7Fit, level = 0.95, vcov. = vcovHC(horvathAge7Fit, type="HC0"))[2,1], 3), ', ', round(coefci(x = horvathAge7Fit, level = 0.95, vcov. = vcovHC(horvathAge7Fit, type="HC0"))[2,2], 3))
EAassoc_age7[1,3] = coeftest(horvathAge7Fit, vcovHC(horvathAge7Fit, type="HC0"))[2,4]

# Skin & bood EAA unadjusted
skinbloodAge7Fit = rlm(DNAmAgeSkinBloodClockAdjAge_age7 ~ DNAmAge_cord_wks, data = datAge7)
EAassoc_age7[1,4] = coeftest(skinbloodAge7Fit, vcovHC(skinbloodAge7Fit, type="HC0"))[2,1]
EAassoc_age7[1,5] = paste0('(', round(coefci(x = skinbloodAge7Fit, level = 0.95, vcov. = vcovHC(skinbloodAge7Fit, type="HC0"))[2,1], 3), ', ', round(coefci(x = skinbloodAge7Fit, level = 0.95, vcov. = vcovHC(skinbloodAge7Fit, type="HC0"))[2,2], 3))
EAassoc_age7[1,6] = coeftest(skinbloodAge7Fit, vcovHC(skinbloodAge7Fit, type="HC0"))[2,4]

# Horvath EAA adjusted
horvathAge7FitAdj = rlm(AgeAccelerationResidual_age7 ~ DNAmAge_cord_wks + female_d + preterm + zvalue_bwt_gage_sex_d + race_child2 + agemom_scale + bmi_scale + coll_grad + smk_preg2 + Bcell_perc_age7 + CD4T_perc_age7 + CD8T_perc_age7 + Neu_perc_age7 + Mono_perc_age7 + NK_perc_age7, data = datAge7)
EAassoc_age7[2,1] = coeftest(horvathAge7FitAdj, vcovHC(horvathAge7FitAdj, type="HC0"))[2,1]
EAassoc_age7[2,2] = paste0('(', round(coefci(x = horvathAge7FitAdj, level = 0.95, vcov. = vcovHC(horvathAge7FitAdj, type="HC0"))[2,1], 3), ', ', round(coefci(x = horvathAge7FitAdj, level = 0.95, vcov. = vcovHC(horvathAge7FitAdj, type="HC0"))[2,2], 3))
EAassoc_age7[2,3] = coeftest(horvathAge7FitAdj, vcovHC(horvathAge7FitAdj, type="HC0"))[2,4]

# Skin & blood EAA adjusted
skinbloodAge3FitAdj = rlm(DNAmAgeSkinBloodClockAdjAge_age7 ~ DNAmAge_cord_wks + female_d + preterm + zvalue_bwt_gage_sex_d + race_child2 + agemom_scale + bmi_scale + coll_grad + smk_preg2 + Bcell_perc_age7 + CD4T_perc_age7 + CD8T_perc_age7 + Neu_perc_age7 + Mono_perc_age7 + NK_perc_age7, data = datAge7)
EAassoc_age7[2,4] = coeftest(skinbloodAge3FitAdj, vcovHC(skinbloodAge3FitAdj, type="HC0"))[2,1]
EAassoc_age7[2,5] = paste0('(', round(coefci(x = skinbloodAge3FitAdj, level = 0.95, vcov. = vcovHC(skinbloodAge3FitAdj, type="HC0"))[2,1], 3), ', ', round(coefci(x = skinbloodAge3FitAdj, level = 0.95, vcov. = vcovHC(skinbloodAge3FitAdj, type="HC0"))[2,2], 3))
EAassoc_age7[2,6] = coeftest(skinbloodAge3FitAdj, vcovHC(skinbloodAge3FitAdj, type="HC0"))[2,4]


# 25 vs 75th percentile difference
predict(horvathAge3Fit, data.frame(DNAmAge_cord_wks = as.vector(quantile(datAge3$DNAmAge_cord_wks, na.rm = T))))
# -0.61612935 -0.24287596 -0.03876924  0.19236200  1.35717599
0.19236200 + 0.24287596
# 0.24287596

predict(horvathAge7Fit, data.frame(DNAmAge_cord_wks = as.vector(quantile(datAge7$DNAmAge_cord_wks, na.rm = T))))
# -1.0257080 -0.4155203 -0.1781412  0.1228073  1.5271482
0.1228073 + 0.4155203
# 0.5383276
```

### Early childhood 
```{r, echo = F}
EAassoc_age3 %>% kable() %>% kable_styling(font_size = 14)
```

### Mid-childhood 
```{r, echo = F}
EAassoc_age7 %>% kable() %>% kable_styling(font_size = 14)
```